---
title: "Gun Control Chi-Square"
author: "Jenny Balmagia"
date: "November 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This: I am Jenny and I am updating this Markdown file here. 

```{r}
library(tidyverse)

gc_df <- read_csv("gc_df.csv")
```
We are asking: is there a significant difference in recommended gun control action for male and female respondents (should guns be easier, harder, or the same difficulty to get?)


Now I (Kyle Monper aka Moo) am adding text:
Hi Pi!

Hi Pi!
The apple I had for lunch was gross. I spit it out and threw it away...

Data Science again...

Jenny: Data wrangling and counts

```{r}
get_guns <- gc_df %>% 
  filter(gender != "NA", make_get_guns != "NA") %>% 
  count(gender, make_get_guns) %>% 
  spread(make_get_guns, n) %>% #takes data from long (tidy) format to make it a contingency table
  select( -gender) #table only contains counts so we can run chi-square
  rownames(get_guns) <- c("female", "male")#assign row names so we can keep track
```

Kyle says: its my turn to do some stats! weeeeee
-- chi-square test for independence



We ask: is there a significant association between sex and opinion on whether guns should be "more dificult" "same" or "less difficult" to get?
```{r}
#Lets look at the actual proportions:

guns_prop <- prop.table(as.matrix(get_guns), 1)


# Run chi-square test for independence:
guns_chi <- chisq.test(get_guns)
guns_chi
```
"Reccommended government action action on gun access differed significantl between male (n= ) and female (n= ) responents" ($\chi^2$(`r guns_chi$paramenter`, p < 0.001))Notably, a much higher proportion on female respondents (57.4%) recommend making it more difficult to get guns, compare that to 40.4% of male respondents.




